---
title: "Analysis of Pilot Study Data"
author: "Rajesh Sigdel"
date: "11/14/2020"
output:
  html_document:
    df_print: paged
---

This is the analysis of Pilot conducted of woody debris at North Buffalo Creek at Greensboro on October 23 and 24, 2020. Three different reaches (sites) of the creek were selected: Latham Park, Audubon Preserve, and Lake Daniel Park.m (see map) Two hundred meters of each reach was surveyed by following Washington State Large Woody Debris Surveys Protocol. Following metrics were evaluated.

* What is the frequency of woody in each site?
* Is the frequency of Lage wood statistically significant?
* What is the volume of Large Wood in each site?
* Are the volumes large Woods of each site significantly different from each other?
* Is the length of Large Wood of each site significantly different from each other?
* Is the length of Large Wood of each site significantly different from each other?




# Importing data
We need to import the dataset first. The dataset can be downloaded from [here](https://drive.google.com/file/d/1JoFpbJSQZwklf41dkDsJufUL0Gd-l6mn/view). (If using windows, press “control” button and [click here](https://drive.google.com/file/d/1JoFpbJSQZwklf41dkDsJufUL0Gd-l6mn/view) it will open in your default browser)


```{r, warning = F, message = F}
library(readxl)

LathamPark <- read_excel("D:/Pilot_Study/Data.xlsx", 
                   sheet = "LathamPark")


LakeDanielPark <- read_excel("D:/Pilot_Study/Data.xlsx", 
                         sheet = "LakeDanielPark")

Auduban <- read_excel("D:/Pilot_Study/Data.xlsx", 
                             sheet = "Auduban")
```


# Installing packages 

```{r, warning = F, message = F}
# install.packages("tidyverse")
library(tidyverse)
# install.packages("devtools")
# devtools::install_github("thomasp85/patchwork")
library(patchwork)
# install.packages("gdata")
library(gdata)
# install.packages("flextable")
library(flextable)
#install.packages("leaflet")
library(leaflet)
```


## Map of the Study site

```{r}
library(readxl)
data <- read_excel("Data_for_map.xlsx")




m <- leaflet() %>% 
  addTiles() %>% 
  addMarkers( data = data, lng = ~Long, lat = ~lat, 
              popup = ~ Name)
m
```



We are now converting all the Null Values into zeros using a function

```{r}
hybrd.ifelse <- function(x) {
  mutate_all(x, ~ifelse(is.na(.), 0, .)) 
}

```

Now, applying the function into our dataframe

```{r}
LathamPark <- hybrd.ifelse(LathamPark)
Auduban <- hybrd.ifelse(Auduban)
LakeDanielPark <- hybrd.ifelse(LakeDanielPark)
```

\

## The formula to calculate Volume is: 

\

```{r}
Volume_Calc <- function(dataframe, 
                   Zone1, Zone2, Zone3, Zone4, PieceD) {
  length <-  (Zone1 + Zone2 + Zone3+ Zone4)
  diameter_m <- (PieceD/100) # Converting diameter into meter
  radius <- diameter_m/2     # Radius = d/2
  volume <- (length*pi*(radius^2))  # L*pi*r^2
  a <- cbind(dataframe, length, diameter_m, radius, volume)
}
```


Applying the function into our dataframe


```{r}
LathamPark <- Volume_Calc(LathamPark, LathamPark$Zone1, LathamPark$Zone2, 
         LathamPark$Zone3, LathamPark$Zone4, LathamPark$PieceD)



Auduban <- Volume_Calc(Auduban, Auduban$Zone1, Auduban$Zone2, 
                         Auduban$Zone3, Auduban$Zone4, Auduban$PieceD)

LakeDanielPark <- Volume_Calc(LakeDanielPark, LakeDanielPark$Zone1, 
                                LakeDanielPark$Zone2, 
                                LakeDanielPark$Zone3, LakeDanielPark$Zone4, 
                                LakeDanielPark$PieceD)
```

Selecting our volume column
```{r, message = F, warning = F}
Auduban_V <- mutate (Auduban, "Auduban Volume" = volume) %>% 
  select("Auduban Volume")
LakeDanielPark_V <- mutate(LakeDanielPark, "Lake Daniel Park volume" = volume) %>% 
  select("Lake Daniel Park volume")
LathamPark_V <- mutate(LathamPark, "Latham Park volume" = volume) %>% 
  select("Latham Park volume")
```


```{r}
Volume <- gdata::cbindX(LakeDanielPark_V, Auduban_V,
       LathamPark_V)
```


```{r}
Volume <- pivot_longer(Volume, c("Lake Daniel Park volume", "Latham Park volume", 
                       "Auduban Volume" ), names_to = "Field",
             values_to = "Volume") %>% drop_na()

```

```{r}
ggplot(Volume, aes(x = Field, y = Volume))+
  geom_boxplot()+
  theme_bw()+
  annotate(geom="text", x=1.3, y=6.5, label="Very Large tree",
           color="red")
```

```{r}
Volume_hist<- function(dataframe, Site){
  a <- filter(dataframe, Field == Site)
  b <- ggplot(a, aes(x = Volume, fill= Site))+
    geom_histogram()+
    labs (x = Site)+
    theme_bw()+
    theme(legend.position = "none")
}
```





```{r}
Auduban_V_ggplot <- Volume_hist(Volume, "Auduban Volume")
LathamPark_V_ggplot <- Volume_hist(Volume, "Latham Park volume")
LakeDanielPark_V_ggplot <- Volume_hist(Volume, "Lake Daniel Park volume")

```



```{r}
All_site_hist <- 
  ggplot(Volume, aes(x = Volume, fill = Volume))+
  geom_histogram()+
  labs (x = "All sites")+
  theme_bw()+
  theme(legend.position = "none")
```


```{r, message = F, warning = F}
(Auduban_V_ggplot | LathamPark_V_ggplot) / 
  (LakeDanielPark_V_ggplot | All_site_hist)
```

The average, median and standard deviation of wood volume in Lake Daniel

```{r}
mean(LakeDanielPark$volume)
sd(LakeDanielPark$volume)
median(LakeDanielPark$volume)
```



The average, median and standard deviation of wood volume in Audubon

```{r}
mean(Auduban$volume)
sd(Auduban$volume)
median(Auduban$volume)
```



The average, median and standard deviation of wood volume in Latham Park

```{r}

mean(LathamPark$volume)
sd(LathamPark$volume)
median(LathamPark$volume)
```


```{r}
a <- Volume %>% 
  group_by(Field) %>%
  summarise(
    "Volume Per 100 meters"  = (sum(Volume))/2,  # Per 100 meter
    mean = mean(Volume, na.rm = TRUE),
    sd = sd(Volume, na.rm = TRUE),
    median = median(Volume, na.rm = TRUE),
    IQR = IQR(Volume, na.rm = TRUE))
```


```{r}
a <- flextable(a)
autofit(a, add_w = 0.001, add_h = 0.1)
```


Jams are excluded in volume calculation because the Washington Survey Guide did not include diameter and length to be calculated for Jams)


# Kruskal Wallis test

Kruskal Wallis test is a non - parametric test alternative to ANOVA. This methos is useful for testing whether sample originates from the same population. The reasons for the test:
•	Volume is a continious data unlike total count of wood logs
•	The dataset is not normal (very skewed data)

```{r}
boxplot(Volume~Field, data = Volume)
```

The result of Kurskal Wallis test is:

```{r}
kruskal.test(Volume ~ Field, data = Volume)
```

There is no statistically significant between volume of individual pieces of Lake Daniel, Lake Latham Park and Auduban Preserve.

---

# Frequency of Large wood per 100 meters

```{r}
Field <- c("Latham Park", "Lake Daniel" , "Auduban")

Values <- c(33, 25, 60)

Pieces <- data.frame(Field, Values)

Pieces$Per100 <- Values/2

ggplot(data = Pieces, aes(x = Field, y = Per100))+
  geom_col(width = 0.5)+
  theme_bw()+
  ylab("Pieces per 100 meters")+
  xlab("")

```


Ho: p1 = p2= p3
Ha: At least one of the site is different from other

```{r}
data <- Pieces$Per100
chisq.test(data)
```
The difference in the frequency of Large Wood is statistically significant

---


# Channel Orientation

```{r}



CountLP <- LathamPark %>% group_by(`Channel Orient`) %>% count(name = "LathamPark") 
CountAu <- Auduban %>% group_by(`Channel Orient`) %>% count(name = "Auduban")
CountLD <- LakeDanielPark %>% group_by(`Channel Orient`) %>% count(name = "LakeDanielPark")





a <- CountLP %>% right_join(CountAu) %>% right_join(CountLD)


flextable::flextable(a)


```



---

# Degradation Class

```{r}
mean(Auduban$`Piece Decay`)

sd(Auduban$`Piece Decay`)

mean(LathamPark$`Piece Decay`)

sd(LathamPark$`Piece Decay`)

mean(LakeDanielPark$`Piece Decay`)

sd(LakeDanielPark$`Piece Decay`)


```


---

# Jams

```{r}

Sites <- c("Auduban Preserve", "Latham Park", "Lake Daniel Park")
`No. of Jams` <- c(2, 1, 0)

data.frame(Sites, `No. of Jams`)

```


---

To find the actual length, we need to add all four zones together. The four zones are: zone 1, zone 2, zone 3, and zone 4. We can use a function to perform the task




```{r, message = F, warning = F}
library(readxl)

LathamPark <- read_excel("D:/Pilot_Study/Data.xlsx", 
                   sheet = "LathamPark")


LakeDanielPark <- read_excel("D:/Pilot_Study/Data.xlsx", 
                         sheet = "LakeDanielPark")

Auduban <- read_excel("D:/Pilot_Study/Data.xlsx", 
                             sheet = "Auduban")

LathamPark <- hybrd.ifelse(LathamPark)
Auduban <- hybrd.ifelse(Auduban)
LakeDanielPark <- hybrd.ifelse(LakeDanielPark)


LathamPark <- Volume_Calc(LathamPark, LathamPark$Zone1, LathamPark$Zone2, 
         LathamPark$Zone3, LathamPark$Zone4, LathamPark$PieceD)



Auduban <- Volume_Calc(Auduban, Auduban$Zone1, Auduban$Zone2, 
                         Auduban$Zone3, Auduban$Zone4, Auduban$PieceD)

LakeDanielPark <- Volume_Calc(LakeDanielPark, LakeDanielPark$Zone1, 
                                LakeDanielPark$Zone2, 
                                LakeDanielPark$Zone3, LakeDanielPark$Zone4, 
                                LakeDanielPark$PieceD)


```

Selecting our Length column
```{r, message = F, warning = F}
Auduban_Length <- mutate (Auduban, "Auduban length" = length) %>% 
  select("Auduban length")
LakeDanielPark_Lenght <- mutate(LakeDanielPark, "Lake Daniel Park length" = length) %>% 
  select("Lake Daniel Park length")
LathamPark_Length <- mutate(LathamPark, "Latham Park length" = length) %>% 
  select("Latham Park length")
```



```{r}
Length <- gdata::cbindX(Auduban_Length, LakeDanielPark_Lenght,
       LathamPark_Length)
```

```{r}
Length <- pivot_longer(Length, c("Lake Daniel Park length", "Latham Park length", 
                       "Auduban length" ), names_to = "Field",
             values_to = "length") %>% drop_na()
```

# Perforning Kruskal Wallis Test on Length 


```{r}
ggplot(data = Length, aes(Field, y = length))+
  geom_boxplot()+
  theme_bw()
```


```{r}
kruskal.test(length ~ Field, data = Length)
```

# Performing Kurskal Wallis Test on Diameter

```{r}
Auduban_Diameter <- Auduban %>% 
  select(diameter_m) %>% 
  rename(diameter_Audobon = diameter_m)

LakeDanielPark_Diameter <- LakeDanielPark %>% 
  select(diameter_m) %>% 
  rename(diameter_LakeDanielPark = diameter_m)


LathamPark_Diameter <- LathamPark %>% 
  select(diameter_m) %>% 
  rename(diamater_LathamPark = diameter_m)
```


```{r}
Diameter <- gdata::cbindX(Auduban_Diameter, LakeDanielPark_Diameter,
       LathamPark_Diameter)
```


```{r}
Diameter <- pivot_longer(Diameter, c("diameter_Audobon", "diameter_LakeDanielPark", 
                       "diamater_LathamPark" ), names_to = "Field",
             values_to = "diameter") %>% drop_na()
```


```{r}
ggplot(data = Diameter, aes(x = diameter, fill = "red"))+
  geom_histogram()+
  theme_bw()+
  theme(legend.position = "none")+
  facet_wrap(~Field)
```


```{r}
kruskal.test(diameter ~ Field, data = Diameter)
```


Feel free to contact me if you have any questions

Thanks

\
\
\
    
  
  